---
title: "Text Analysis: Regular Expressions"
subtitle: "<br><br> IS 407"
author: "JooYoung Seo"
output:
  xaringan::moon_reader:
    css: ["../xaringan-themer.css", "../slides.css"]
    lib_dir: libs
    anchor_sections: FALSE
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "../setup.Rmd"}
```

class: middle

# Regular Expression (Regex)

---

## Regex

- Represented as a single string and defines a set of matching strings.
- Learn more at [RStudio regex cheatsheets](https://www.rstudio.com/resources/cheatsheets#regular-expressions).

---

## Nitty-gritty

- `[character]{min,max}`

- `[ ]`: possible character choices

- `{ , }`: number of repitition (or range) for the character

- `^`: start of string (ctrl+home)

- `$`: end of string (ctrl+end)

---

## Example

```{r}
library(tidyverse)

text <- c("is407", "a", "H2O", "cat", "Cat", "cats", "the cat", "123", "111", "11111", "jonathan", "johnathan") #<<
df <- tibble(text)

df
```

---

## Filter cat?

```{r}
df %>%
  filter(
    str_detect(
      text,
      "cat" #<<
    )
  )
```
---

## Filter only cat

### Exclude "the cat"

```{r}
df %>%
  filter(
    str_detect(
      text,
      "^cat" #<<
    )
  )
```

---

## Filter only cat

### Exclude plural

```{r}
df %>%
  filter(
    str_detect(
      text,
      "^cat$" #<<
    )
  )
```

---

## Filter cat and Cat

```{r}
df %>%
  filter(
    str_detect(
      text,
      "^[cC]at$" #<<
    )
  )
```

---

## Filter all the alphabetic characters?

```{r eval = FALSE}
df %>%
  filter(
    str_detect(
      text,
      "[]" #<<
    )
  )
```

---

## Filter all the alphabetic characters?

```{r}
df %>%
  filter(
    str_detect(
      text,
      "[abcdefghijklmnopqrstuvwxyzABCDEFGHIJKLMNOPQRSTUVWXYZ]" #<<
    )
  )
```

---

## Filter all the alphabetic characters?

### Shorten using dash

```{r}
df %>%
  filter(
    str_detect(
      text,
      "[a-zA-Z]" #<<
    )
  )
```

---

## Filter all the alphabetic characters?

### Don't want to include "is407"

```{r}
df %>%
  filter(
    str_detect(
      text,
      "[a-zA-Z]$" #<<
    )
  )
```


---

## Filter all the alphabetic characters?

### Don't want to include H2O

```{r}
df %>%
  filter(
    str_detect(
      text,
      "^[a-zA-Z]$" #<<
    )
  )
```

---

## Filter all the alphabetic characters?

### the previous slide is the same as this

```{r}
df %>%
  filter(
    str_detect(
      text,
      "^[a-zA-Z]{1}$" #<<
    )
  )
```



---

## Filter all the alphabetic characters?

### How many alphabetic patterns?: `{min,max}`

```{r}
df %>%
  filter(
    str_detect(
      text,
      "^[a-zA-Z]{1,3}$" #<<
    )
  )
```

---

## Filter all the alphabetic characters?

### For unknown/unlimited max: `{min,}`

```{r}
df %>%
  filter(
    str_detect(
      text,
      "^[a-zA-Z]{1,}$" #<<
    )
  ) %>%
  print(n = 12)
```

---

## Bonus: shorthand class
 
### POSIX-compliant regex

- `[[:alpha:]]` is the same as `[a-zA-Z]`

- `+` is the same as `{1,}`

- `[[:alpha:]]+` == `[a-zA-Z]{1,}`


---

## Filter all the alphabetic characters?

### Shorthand form

```{r}
df %>%
  filter(
    str_detect(
      text,
      "^[[:alpha:]]+$" #<<
    )
  ) %>%
  print(n = 12)
```

---

.question[
# Quizzes

1. Find "11111".

2. Find digit(s)-only words.

3. Find both "jonathan" and "johnathan".

4. Find "is407" and "H2O" using regex pattern.
]

---

## Find "11111"

```{r}
df %>%
  filter(
    str_detect(
      text,
      "[1]{5}" #<<
    )
  ) %>%
  print(n = 12)
```

---

## Find digits only

```{r}
df %>%
  filter(
    str_detect(
      text,
      "^[0123456789]{1,}$" #<<
    )
  ) %>%
  print(n = 12)
```

---

## Find digits only

```{r}
df %>%
  filter(
    str_detect(
      text,
      "^[0-9]{1,}$" #<<
    )
  ) %>%
  print(n = 12)
```

---

## Bonus

- `[[:digit:]]` == `[0-9]`

- `\\d` == `[0-9]`

---

## Find digits only: shorthand v1

```{r}
df %>%
  filter(
    str_detect(
      text,
      "^[[:digit:]]+$" #<<
    )
  ) %>%
  print(n = 12)
```


---

## Find digits only: shorthand v2

```{r}
df %>%
  filter(
    str_detect(
      text,
      "^\\d+$" #<<
    )
  ) %>%
  print(n = 12)
```

---

## Find "jonathan" and "johnathan"

```{r}
df %>%
  filter(
    str_detect(
      text,
      "jo[h]{0,1}nathan" #<<
    )
  ) %>%
  print(n = 12)
```

---

## Bonus

- `h` == `[h]`

    - because there is only one choice.

- `?` == `{0,1}`

- `[h]{0,1}` == `h?`

---

## Find "jonathan" and "johnathan"

### Shorthand form

```{r}
df %>%
  filter(
    str_detect(
      text,
      "joh?nathan" #<<
    )
  ) %>%
  print(n = 12)
```

---

## Find "is407" and "H2O"

```{r}
df %>%
  filter(
    str_detect(
      text,
      "[a-zA-Z]{1,2}[0-9]{1,3}[a-zA-Z]{0,1}" #<<
    )
  ) %>%
  print(n = 12)
```

---

## Find "is407" and "H2O": shorthand version

```{r}
df %>%
  filter(
    str_detect(
      text,
      "[[:alpha:]]{1,2}\\d{1,3}[[:alpha:]]?" #<<
    )
  ) %>%
  print(n = 12)
```

---

## Learn more about Regex

- [RStudio regex cheatsheets](https://www.rstudio.com/resources/cheatsheets#regular-expressions).
